### 2023-01-19 extras/mltl_in_the_wild.tex
formulas in the wild

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsthm, amsmath, amssymb}
\usepackage[letterpaper,width=150mm,top=20mm,bottom=25mm]{geometry}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{ dsfont }
\usepackage{ mathrsfs }
\usepackage{hyperref}
\usepackage{cancel}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
}

\newcommand{\FF}{\mathcal{F}}
\newcommand{\GG}{\mathcal{G}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\RR}{\mathcal{R}}

\begin{document}
From \href{https://link.springer.com/chapter/10.1007/978-3-030-76384-8_10}{Integrating Runtime Verification into a Sounding Rocket Control System}:
\begin{itemize}
    \item $\neg \GG_{[0,2]} \neg p_0$
    \item $p_0 \rightarrow \FF_{[0,140]}p_1$
    \item $p_0 \rightarrow \FF_{[0,800]}p_1$
    \item $(p_0 \land p_1) \rightarrow \FF_{[0,10]}p_2$
    \item $p_0 \rightarrow (p_0 \UU_{[0,130]} p_1)$
    \item $p_0 \rightarrow \FF_{[0,5]} p_1$
    \item $(p_0 \land p_1) \rightarrow \FF_{[0, 126]} p_2$
    \item $p_0 \rightarrow (p_0 \UU_{[0,114]} p_1)$
\end{itemize}
$$$$

From \href{https://link.springer.com/chapter/10.1007/978-3-030-57628-8_12}{Embedding Online Runtime Verification for Fault Disambiguation on Robonaut2}:
\begin{itemize}
    \item $\GG_{[0,3]}p_0$
    \item $(p_0 \land \GG_{[0,3]}p_1) \rightarrow p_2$
\end{itemize}
$$$$

From \href{https://link.springer.com/chapter/10.1007/978-3-031-06773-0_45}{
Runtime verification triggers real-time, autonomous fault recovery on the CySat-I}:
\begin{itemize}
    \item $\GG_{[0,360]}(\neg p_0 \land \neg p_1 \land \neg p_2 \land \neg p_3 \land \neg p_4 \land \neg p_5)$
    \item $\GG_{[360, M]}(p_0 \rightarrow p_1)$ ($M$ is end of mission)
    \item $\GG_{[0, M]}(p_0 = p_1)$
\end{itemize}
$$$$

From \href{https://link.springer.com/chapter/10.1007/978-3-030-59155-7_26}{Integrating Runtime Verification into an Automated UAS Traffic Management System}:
\begin{itemize}
    \item $\neg \GG_{[0,3]} \neg (p_0 \land p_1)$
    \item $\GG_{[0,3]} (p_0 \land p_1)$
    \item $(p_0 \land p_1 \land p_2) \rightarrow \neg \GG_{[0,3]}\neg(p_3 \land p_4)$
    \item $\GG_{[0,3]} \neg(p_0 \land p_1)$
    \item $\GG_{[0, M]} (p_0 \lor p_1 \lor p_2 \lor p_3 \lor p_4 \lor p_5 \lor p_6 \lor p_7 \lor p_8)$ ($M$ is end of mission)
    \item $\GG_{[0, M]}(p_0 \rightarrow (p_1 \land p_2))$
    \item $\GG_{[0,M]}((p_0 \lor p_1) \rightarrow (p_2 \land p_3 \land p_4 \land p_5 \land p_6))$
    \item $p_0 \rightarrow \FF_{[0,M]} p_1$
    \item $p_0 \rightarrow \FF_{[0,M]} (p_1 \land p_2 \land p_3)$
    \item $\FF_{[0,M]} (p_1 \land p_2)$
\end{itemize}
$$$$

From \href{https://link.springer.com/chapter/10.1007/978-3-030-25543-5_1}{
Satisfiability Checking for Mission-Time LTL}:\\
\indent Consider random mLTL formula generation for benchmarking\\

\end{document}